Commutative character sheaves and geometric types for supercuspidal representations

### 2017-08-30 Geometric types.tex
Oops. Pushed the wrong version last time. Here we go ...

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.}
\end{quotation}
In this paper we follow the suggestion above by showing that certain Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence of representations $\orho_i$ of compact $p$-adic groups $\oK^i$, for $i=0, \ldots, d$, such that $(\oK^d, \rho_d)$ is a type for a supercuspidal representation of a $p$-adic group.
Let $R$ be the ring of integers of a local field with finite residue field $k$.
The main result of \cite{Yu:models} shows how to find, for each $i=0, \ldots, d$, a smooth group scheme $\underline{G}^i$ over the ring $R$ with $\underline{G}^i(R)=\,^\circ K^i $.
The geometrization of the Yu type datum that we consider in this paper uses Lusztig's theory of character sheaves on reductive groups over finite fields, for which it is necessary to assume that the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic. 
We must also place a condition on the linear characters that appear in the Yu datum. 
Under these assumptions we show how each representation $\orho_i$ can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, where $\cs{F}^i$ is a {\it rational virtual} sheaf complex on the Greenberg transform $G^i$ of $\underline{G}^i$, by which we mean $\cs{F}^i$ is an element in the Grothendieck group of such sheaves, tensored with $\QQ$. 
Writing $\trFrob{\cs{F}^i}$ for the function on $G^i(k) = \underline{G}^i(R) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the rational virtual complex $\cs{F}^i$, we show in Theorem~\ref{thm:geotypes} that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i).
\end{equation}
By this theorem, then, we obtain geometric avatars for each type in a Yu datum:
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob}}
(\underline{G}^i, \cs{F}^i).
\end{tikzcd}
\]
We refer to the pair $(\underline{G}^d, \cs{F}^d)$ as a \emph{geometric type}.

To prove Theorem~\ref{thm:geotypes}, we must find a way to geometrize linear characters of groups of the form $\underline{H}(R)$, where $\underline{H}$ is a smooth group scheme over $R$. 
In order to do so in a systematic manner, we begin this paper by describing the function-sheaf dictionary for certain \todo{Certain?} characters of arbitrary smooth group schemes over finite fields. 
When coupled with the Greenberg transform, this dictionary will allow for the geometrization of linear characters of $\underline{H}(R)$.

@@

In order to provide further justification for referring to objects in $\CCS(G)$ as commutative character sheaves, suppose for the moment that $G$ is a connected, reductive algebraic group over $\Fq$.
Let $\gcs{L}$ be the geometric part of an object in $\CCS(G)$; see Section~\ref{sec:defs}.
Let $T$ be a maximal torus in $\bG$ and let $\gcs{L}_T$ be the restriction of $\gcs{L}$ to $T$.
Then the perverse sheaf $\gcs{L}[\dim G]$ appears in the semisimple complex $\operatorname{ind}_{B,T}^{\bG}(\gcs{L}_T)$ produced by parabolic induction.
It follows that every object in $\CCS(G)$ determines a Frobenius-stable character sheaf on $G$, in the sense of \cite{lusztig:85a}*{Def.~2.10}. \todo{Must revisit this}
Of course, the sheaves arising in this way represent a small part of Lusztig's geometrization of characters of representations of connected, reductive groups over finite fields, but they are precisely those needed to describe one-dimensional characters of such groups.

@@

\begin{proposition}\label{prop:quasicharacters}
Let $\underline{G}$ be a smooth group scheme over $R$.
Let $\varphi : \underline{G}(R) \to \EEx$ be quasicharacter and let $r$ be the level of $\varphi$.
If $\varphi_r : G_m(k)\to \EEx$ is trivial on $G_{r,\operatorname{der}}(k)$ then 
there is a commutative character sheaf $\cs{L}_r\in \CCS(G_r)$ such that 
\[
\trFrob{\cs{L}_r} =  \varphi_r.
\] 
\end{proposition}

@@

Recall that the full Greenberg transform $G \ceq \Gr^{R}(\underline{G})$ is a group scheme over $\Fq$ such that $G(\Fq) = \underline{G}(R)$; it comes equipped with a morphism $G \to G_m$.
The Weil sheaf on $G$ obtained from $\cs{L}_m$ by pullback along the morphism of group schemes $G \to G_m$ is a quasicharacter sheaf on $G$, in the sense of \cite{cunningham-roe:13a}*{Def 4.2}, such that 
\[
\trFrob{\cs{L}} = \varphi.
\]

@@

For the rest of Section~\ref{sec:types}, $K$ is a $p$-adic field and $R$ is the ring of integers of $K$.
A Yu type datum consists of the following:
\begin{enumerate}
\labitem{Y0}{Y0} a sequence of compact groups $\oK^0 \subseteq \oK^1 \subseteq \cdots \subseteq \oK^d = \oK$;
\labitem{Y1}{Y1} a continuous representation $\orho^0$ of $\oK^0$;
\labitem{Y2}{Y2} quasicharacters $\varphi^i : \oK^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
The representation $\orho^0$ and the quasicharacters $(\varphi^0, \ldots , \varphi^d)$ enjoy certain properties which allow Yu to construct the representations $\orho_i$ of $\oK^i$ that form the sequence of types $(\oK^i,\orho_i)$, for $i=1, \ldots, d$.
%
In order to prepare for the construction of the geometric types of Theorem~\ref{thm:geotypes}  we review some further detail here.
In Table~\ref{table:notation} we explain how to convert the constructions appearing in this section into the notation of \cite{yu:01a}.

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.3}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho^{i+1}$ & $\,^\circ \rho^{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \cite{yu:01a}*{\S 3}\\
$J_{i+1}$ & $J^{i+1} = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ & \cite{yu:01a}*{\S 3} \\
$V_{i+1}$ & $J^{i+1}/J^{i+1}_+ = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \cite{yu:01a}*{\S 3} \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ &  \cite{yu:01a}*{\S 4} \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%

We can now recall how Yu uses all this to construct representations $\orho^i$ of $\oK^i$, for $i=1, \ldots, d$ and the types $(\oK^i,\orho_i)$; see \cite{yu:01a}*{\S\S 4, 15}.
The representations $\orho^i$ and $\orho_i$ are defined recursively.
For the base case $i=0$, set $\orho_0 \ceq \orho^0\otimes \varphi^0$; see \ref{Y1} above.
Now fix $i$.
Let $W_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$, whose restriction to $V_{i+1}^\sharp$ has central character $\psi_{i+1}$.
Pull-back along $h_{i+1}$ to form $h_{i+1}^*(W_{i+1})$, a representation of $\oK^i \ltimes J_{i+1}$.
Write $\inf(\orho_i)$ for the representation of $\oK^i \ltimes J_{i+1}$ obtained by pulling back $\orho^i$ along $\oK^i \ltimes J_{i+1} \to \oK^i$. 
Consider the representation
\begin{equation}\label{eq:tensor}
\orho^{i+1} \ceq h_{i+1}^*(W_{i+1}) \otimes \inf(\orho_i)
\end{equation}
of $\oK^i \ltimes J_{i+1}$.
By \cite{yu:01a}*{}, the representation $\orho^{i+1}$ of $\oK^i \ltimes J_{i+1}$ is trivial on $\oK^{i}\cap J_{i+1}$ so $\orho^{i+1}$ descends to $\oK^{i+1}$. 
%
Set $\orho_{i+1} = \orho^{i+1}\otimes \varphi^{i+1}$.
This completes the recursive definition of the Yu $(\oK^i,\orho_i)$ for $i=0, \ldots , d$.
By \cite{Yu:models}*{Prop 10.2} there is a sequence  
\[
\underline{G}^0 \to \underline{G}^1 \to \cdots \to \underline{G}^d = \underline{G}
\]
of morphisms of affine smooth group schemes of finite type over $R$ such that, on $R$-points it gives the sequence $\oK^0 \subseteq \oK^1 \subseteq \cdots \subseteq \oK^d$ above.
Indeed, this is the main result of \cite{Yu:models}.

@@

We may now revisit the ingredients in the construction of the representation $\rho$ of $\underline{G}(R)$ along the lines indicated by Yu and recalled in Section~\ref{ssec:review}.
\begin{enumerate}
\labitem{M0}{M0}
The compact groups $\oK^i$ appearing in \ref{Y0} have been replaced by the smooth group schemes $\underline{G}^i$.
\labitem{M1}{M1}
The continuous representation $\orho^0$ of $\oK^0$ appearing in \ref{Y1} may be interpreted as a representation of $\underline{G}^0(R)$ obtained by inflation along $\underline{G}^0(R) \to \underline{G}^0(\Fq)$ from a representation $\varrho_0$ of $\underline{G}^0(\Fq) = \underline{G}^0_\Fq(\Fq)$.
In fact, $\varrho_0$ is itself obtained by pulling back a representation $\varrho_0^\reductive$ along the $\Fq$-points of the quotient $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.
\labitem{M2}{M2} The quasicharacters $\varphi^i$ appearing in \ref{Y2} are now quasicharacters of $\underline{G}^i(R)$, for $i=0, \ldots, d$.
\labitem{M3}{M3}
Diagram \eqref{eq:semiprod} in \ref{Y3} is now replaced by the following diagram of smooth group schemes over $R$.
\begin{equation}\label{eq:pimodel}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && \underline{G}^{i}\times_{\underline{G}} \underline{J}^{i+1}\arrow{d} && \\
1 \arrow{r} & \underline{J}^{i+1} \arrow{r} & \underline{G}^i \ltimes \underline{J}^{i+1} \arrow{d} \arrow{r} & \underline{G}^i \arrow{r} & 1\\
&& \underline{G}^{i+1} \arrow{d} && \\
&& 1 &&
\end{tikzcd}
\end{equation}
\labitem{M4}{M4}
The representation $h_{i+1}^*(W_{i+1})$ appearing in \ref{Y4} is now obtained by pulling back a representation along 
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(R) \to (\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq).
\]
Let $w_{i+1}$ be that representation of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq) = (\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq)(\Fq)$. 
Then $w_{i+1}$ is itself obtained by pulling back the representation $W_{i+1}$ along the $\Fq$-points of the quotient
\[
\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq \to 
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp.
\]
\end{enumerate}

@@

We may now give the main result of Section~\ref{sec:types}.
%
Since Yu's theory refers to complex representations, and since our geometrization uses $\ell$-adic sheaves, we grit our teeth and fix an isomorphism $\CC \approx \EE$.
From Section~\ref{ssec:equivariant}, recall the definition of a rational virtual equivariant Weil sheaf complex.

\begin{theorem}\label{thm:geotypes}
Let $\oK^i$ $\orho^0$, $\varphi^i$, for $i=0, \ldots d$, be a Yu type datum as in Section~\ref{ssec:review}, \ref{Y0}, \ref{Y1} and \ref{Y2}.
Let $\underline{G}^i$ and $\varrho_0\red$ be as in Section~\ref{ssec:review}, \ref{M0} and \ref{M1}.
Assume $\pi_0((\underline{G}^0)_\Fq^\reductive)$ is cyclic.
For $i=0, \ldots ,d$, let $G^i = \Gr_{R}(\underline{G}^i)$ be the Greenberg transform of the smooth group scheme $\underline{G}^i$ appearing in Section~\ref{ssec:review}.
Suppose that the quasicharacters $\varphi^i$, given in Section~\ref{ssec:review}, \ref{M2}, satisfy the hypothesis of Proposition~\ref{prop:quasicharacters}.
Then there is a rational virtual equivariant Weil sheaf complex $\cs{F}_i$ on $G^i$ such that $\trFrob{\cs{F}_i} = \trace(\orho_i)$, for $i=0, \ldots , d$.
\end{theorem}

\begin{proof}
%For every $i=0, \ldots ,d$, set $G^i = \Gr_{R}(\underline{G}^i)$.
Recall that $G^i(\Fq) = \underline{G}^i(R) = \oK^i$, canonically.
Set $G_m^i = \Gr_m^{R}(\underline{G}^i)$, for $i=0, \ldots, d$.
Then $G_m^i$ is a smooth group scheme over $\Fq$ and $G_m^i(\Fq) = \underline{G}^i(R_m)$, canonically.
Let $m_i$ be the level of $\varphi$ as defined in Section~\ref{ssec:quasicharacters}.
Let $\cs{L}^i_m$ be the commutative character sheaf on $G_{m_i}$ given by Proposition~\ref{prop:quasicharacters}; then
$\trFrob{\cs{L}^i_{m_i}} = \varphi^i_{m_i}$.
Set $m = \max\{ m_i \tq i=0, \ldots ,d\}$ and, for each $i$, let $\cs{L}^i_{m}$ be the pull-back of $\cs{L}^i_{m_i}$ along $G^i_{m} \to G^i_{m_i}$; then $\cs{L}^i_{m}$ is a linear character sheaf and
\[
\trFrob{\cs{L}^i_{m}} = \varphi^i_{m}.
\]

>>>>

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.}
\end{quotation}

In this paper we follow the suggestion above by showing that certain Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence of representations $\orho_i$ of compact $p$-adic groups $\oK^i$, for $i=0, \ldots, d$, such that $(\oK^d, \rho_d)$ is a type for a supercuspidal representation of a $p$-adic group.
Let $R$ be the ring of integers of a local field with finite residue field $k$.
The main result of \cite{Yu:models} shows how to find, for each $i=0, \ldots, d$, a smooth group scheme $\underline{G}^i$ over the ring $R$ with $\underline{G}^i(R)=\,^\circ K^i $.
Under certain assumptions on the Yu type datum, we show how each representation $\orho_i$ can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, where $\cs{F}^i$ is a formal ${\bar\QQ}$-linear conbination of conjugation equivariant sheaf complexes on the Greenberg transform $G^i$ of $\underline{G}^i$.

Writing $\trFrob{\cs{F}^i}$ for the function on $G^i(k) = \underline{G}^i(R) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the rational virtual complex $\cs{F}^i$, we show in Theorem~\ref{thm:geotypes} that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i).
\end{equation}
By this theorem, then, we obtain geometric avatars for each type in a Yu datum:
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob}}
(\underline{G}^i, \cs{F}^i).
\end{tikzcd}
\]
We refer to the pair $(\underline{G}^d, \cs{F}^d)$ as a \emph{geometric type}.

To prove Theorem~\ref{thm:geotypes}, we must find a way to geometrize linear characters of groups of the form $\underline{H}(R)$, where $\underline{H}$ is a smooth group scheme over $R$. 
In order to do so in a systematic manner, we begin this paper by describing the function-sheaf dictionary for characters of arbitrary smooth group schemes over finite fields. 
When coupled with the Greenberg transform, this dictionary will allow for the geometrization of certain linear characters of $\underline{H}(R)$.

@@

In order to provide further justification for referring to objects in $\CCS(G)$ as commutative character sheaves, suppose for the moment that $G$ is a connected, reductive algebraic group over $\Fq$.
Let $\gcs{L}$ be the geometric part of an object in $\CCS(G)$; see Section~\ref{sec:defs}.
Let $T$ be a maximal torus in $\bG$ and let $\gcs{L}_T$ be the restriction of $\gcs{L}$ to $T$.
Then the perverse sheaf $\gcs{L}[\dim G]$ appears in the semisimple complex $\operatorname{ind}_{B,T}^{\bG}(\gcs{L}_T)$ produced by parabolic induction.
It follows that every object in $\CCS(G)$ determines a Frobenius-stable character sheaf on $G$, in the sense of \cite{lusztig:85a}*{Def.~2.10}. 
Of course, the sheaves arising in this way represent a small part of Lusztig's geometrization of characters of representations of connected, reductive groups over finite fields, but they are precisely those needed to describe one-dimensional characters of such groups.

@@

\begin{proposition}\label{prop:quasicharacters}
Let $\underline{G}$ be a smooth group scheme over $R$.
Let $\varphi : \underline{G}(R) \to \EEx$ be quasicharacter and let $r$ be the level of $\varphi$.
If $\varphi_r : G_r(k)\to \EEx$ is trivial on $G_{r,\operatorname{der}}(k)$ then 
there is a commutative character sheaf $\cs{L}_r\in \CCS(G_r)$ such that 
\[
\trFrob{\cs{L}_r} =  \varphi_r.
\] 
\end{proposition}

@@

Recall that the full Greenberg transform $G \ceq \Gr^{R}(\underline{G})$ is a group scheme over $\Fq$ such that $G(\Fq) = \underline{G}(R)$; it comes equipped with a morphism $G \to G_r$.
The Weil sheaf on $G$ obtained from $\cs{L}_m$ by pullback along the morphism of group schemes $G \to G_r$ is a quasicharacter sheaf on $G$, in the sense of \cite{cunningham-roe:13a}*{Def 4.2}, such that 
\[
\trFrob{\cs{L}} = \varphi.
\]

@@

For the rest of Section~\ref{sec:types}, $K$ is a $p$-adic field and $R$ is the ring of integers of $K$.
A Yu type datum consists of the following:
\begin{enumerate}
\labitem{Y0}{Y0} a sequence of compact groups $\oK^0 \subseteq \oK^1 \subseteq \cdots \subseteq \oK^d = \oK$;
\labitem{Y1}{Y1} a continuous representation $\orho^0$ of $\oK^0$;
\labitem{Y2}{Y2} quasicharacters $\varphi^i : \oK^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
The representation $\orho^0$ and the quasicharacters $(\varphi^0, \ldots , \varphi^d)$ enjoy certain properties which allow Yu to construct the representations $\orho^i$ of $\oK^i$ that form the sequence of types $(\oK^i,\orho_i)$, for $i=1, \ldots, d$.
%
In order to prepare for the construction of the geometric types of Theorem~\ref{thm:geotypes}  we review some further detail here.
In Table~\ref{table:notation} we explain how to convert the constructions appearing in this section into the notation of \cite{yu:01a}.

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.3}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho_{i+1}$ & $\,^\circ \rho_{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \cite{yu:01a}*{\S 3}\\
$J_{i+1}$ & $J^{i+1} = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ & \cite{yu:01a}*{\S 3} \\
$V_{i+1}$ & $J^{i+1}/J^{i+1}_+ = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \cite{yu:01a}*{\S 3} \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ &  \cite{yu:01a}*{\S 4} \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%

We can now recall how Yu uses all this to construct the types $(\oK^i,\orho_i)$; see \cite{yu:01a}*{\S\S 4, 15}.
The representations $\orho_i$ are defined recursively.
For the base case $i=0$, set $\orho_0 \ceq \orho^0\otimes \varphi^0$; see \ref{Y1} above.
Now fix $i$.
Let $W_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$, whose restriction to $V_{i+1}^\sharp$ has central character $\psi_{i+1}$.
Pull-back along $h_{i+1}$ to form $h_{i+1}^*(W_{i+1})$, a representation of $\oK^i \ltimes J_{i+1}$.
Write $\inf(\orho_i)$ for the representation of $\oK^i \ltimes J_{i+1}$ obtained by pulling back $\orho_i$ along $\oK^i \ltimes J_{i+1} \to \oK^i$. 
Consider the representation
\begin{equation}\label{eq:tensor}
\orho^{i+1} \ceq h_{i+1}^*(W_{i+1}) \otimes \inf(\orho_i)
\end{equation}
of $\oK^i \ltimes J_{i+1}$.
By \cite{yu:01a}*{}, the representation $\orho^{i+1}$ of $\oK^i \ltimes J_{i+1}$ is trivial on $\oK^{i}\cap J_{i+1}$ so $\orho^{i+1}$ descends to $\oK^{i+1}$. 
%
Set $\orho_{i+1} = \orho^{i+1}\otimes \varphi^{i+1}$.
This completes the recursive definition of the Yu $(\oK^i,\orho_i)$ for $i=0, \ldots , d$.
By \cite{Yu:models}*{Prop 10.2} there is a sequence  
\[
\underline{G}^0 \to \underline{G}^1 \to \cdots \to \underline{G}^d = \underline{G}
\]
of morphisms of affine smooth group schemes of finite type over $R$ such that, on $R$-points it gives the sequence $\oK^0 \subseteq \oK^1 \subseteq \cdots \subseteq \oK^d$ above.
Indeed, this is the main result of \cite{Yu:models}.

@@

We may now revisit the ingredients in the construction of the representation $\rho$ of $\underline{G}(R)$ along the lines indicated by Yu and recalled in Section~\ref{ssec:review}.
\begin{enumerate}
\labitem{M0}{M0}
The compact groups $\oK^i$ appearing in \ref{Y0} have been replaced by the smooth group schemes $\underline{G}^i$.
\labitem{M1}{M1}
The continuous representation $\orho^0$ of $\oK^0$ appearing in \ref{Y1} may be interpreted as a representation of $\underline{G}^0(R)$ obtained by inflation along $\underline{G}^0(R) \to \underline{G}^0(\Fq)$ from a representation $\varrho_0$ of $\underline{G}^0(\Fq) = \underline{G}^0_\Fq(\Fq)$.
In fact, $\varrho_0$ is itself obtained by pulling back a representation $\varrho_0^\reductive$ along the $\Fq$-points of the quotient $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.
\labitem{M2}{M2} The quasicharacters $\varphi^i$ appearing in \ref{Y2} are now quasicharacters of $\underline{G}^i(R)$, for $i=0, \ldots, d$.
In fact, if $r_i$ is the level of $\varphi^i$, as it appears in Section~\ref{ssec:quasicharacters}, then $\varphi^i$ is obtained by pulling back a quasicharacter $\varphi_i\ceq \varphi^i_{r_i}$ of the smooth group scheme $G_i\ceq G^i_{r_i} \ceq \Gr_{r_i}^{R}(\underline{G}^i)$ along $G^i \to G_i$.
\labitem{M3}{M3}
Diagram \eqref{eq:semiprod} in \ref{Y3} is now replaced by the following diagram of smooth group schemes over $R$.
\begin{equation}\label{eq:pimodel}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && \underline{G}^{i}\times_{\underline{G}} \underline{J}^{i+1}\arrow{d} && \\
1 \arrow{r} & \underline{J}^{i+1} \arrow{r} & \underline{G}^i \ltimes \underline{J}^{i+1} \arrow{d} \arrow{r} & \underline{G}^i \arrow{r} & 1\\
&& \underline{G}^{i+1} \arrow{d} && \\
&& 1 &&
\end{tikzcd}
\end{equation}
\labitem{M4}{M4}
The representation $h_{i+1}^*(W_{i+1})$ appearing in \ref{Y4} is now obtained by pulling back a representation along 
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(R) \to (\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq).
\]
Let $w_{i+1}$ be that representation of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq) = (\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq)(\Fq)$. 
Then $w_{i+1}$ is itself obtained by pulling back the representation $W_{i+1}$ along the $\Fq$-points of the quotient
\[
\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq \to 
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp.
\]
\end{enumerate}

@@

We may now give the main result of Section~\ref{sec:types}, Theorem~\ref{thm:geotypes}.
Since Yu's theory refers to complex representations, and since our geometrization uses $\ell$-adic sheaves, we grit our teeth and fix an isomorphism $\CC \approx \EE$.

As we recalled in Section~\ref{ssec:review}, a Yu type datum consists of compact groups $\oK^i$, a representation $\orho^0$ of $\oK^0$ and quasicharacters $\varphi^i$ of $\oK^i$, for $i=0,\ldots, d$; see \ref{Y0}, \ref{Y1} and \ref{Y2}.
In Section~\ref{ssec:review} we also saw how a Yu type datum determines smooth group schemes $\underline{G}^i$, a representation $\varrho_0\red$ of the reductive quotient $(\underline{G}^0)\red_{k}$ of the special fibre $\underline{G}^0_{k}$ of $\underline{G}^0$, and quasicharacters $\varphi_i$ of the group of $k$-rational points on the level-$r_i$ Greenberg transform $G_i$ of $\underline{G}^i$; see \ref{M0}, \ref{M1} and \ref{M2}.

Theorem~\ref{thm:geotypes} places the following conditions on the Yu type datum:
\begin{enumerate}
\labitem{H0}{H0} the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic;
\labitem{H1}{H1} for each $i=0,\ldots, d$, either
\begin{enumerate}
\labitem{H1(a)}{H1a} $G_i$ is reductive with cyclic geometric component group, or
\labitem{H1(b)}{H1b} the quasicharacter $\varphi_i$ of $G_i$ is trivial on  $G_{i,\operatorname{der}}(k)$.
\end{enumerate}
\end{enumerate}
Hypothesis~\ref{H0} will allow us to use Lusztig's character sheaves to geometrize the character of the representation of $\orho^0$.
Hypothesis~\ref{H1a} allows us to use Proposition~\ref{prop:quasicharacters} to geometrize each quasicharacter $\varphi^i$ using local systems on $G_i$, or more precisely, using commutative character sheaves on $G_i$;
alternatively, Hypothesis~\ref{H1b} may be used to geometrize each quasicharacter $\varphi^i$ using an element in the Grothendieck group of perverse sheaves on $G_i$, tensored with ${\bar\QQ}$.

\begin{theorem}\label{thm:geotypes}
Let $\oK^i$, $\orho^0$ and $\varphi^i$, be a Yu type datum, with $i=0,\ldots, d$, satisfying Hypotheses~\ref{H0} and \ref{H1}.
There is an element $\cs{F}_i$ in the Grothendieck group of conjugation equivariant Weil sheaf complexes on $G^i$, tensored with ${\bar\QQ}$, such that $\trFrob{\cs{F}_i} = \trace(\orho_i)$, where $\orho_i$ is the representation of $\oK^i$ constructed from the Yu type datum in Section~\ref{ssec:review}, for $i=0,\ldots, d$.
\end{theorem}

\begin{proof}
%For every $i=0, \ldots ,d$, set $G^i = \Gr_{R}(\underline{G}^i)$.
Recall that $G^i(\Fq) = \underline{G}^i(R) = \oK^i$, canonically.
Let $r_i$ be the level of $\varphi^i$ as defined in Section~\ref{ssec:quasicharacters}.
Set $r = \max\{ r_i \tq i=0, \ldots ,d\}$.

By \cite{lusztig:disconnected1}, there is a rational virtual Weil character sheaf $A = ({\bar A},\phi)$ on $(\underline{G}^0)_\Fq^\reductive$ such that 
%${\bar A}$ is a rational virtual character sheaf on $(\underline{G}^0)_{\bFq}^\reductive$ and
\[
\trFrob{A} = \trace \varrho_0^\reductive.
\]
This uses Hypothesis~\ref{H0}.
%
Let $A^0$ be the conjugation equivariant Weil sheaf on $(\underline{G}^0)_\Fq$ obtained by pullback along the quotient $(\underline{G}^0)_\Fq \to (\underline{G}^0)_\Fq^\reductive$.
Then 
\[
\trFrob{A^0} = \trace \varrho_0.
\]
The special fibre $(\underline{G}^0)_\Fq$ of the smooth group scheme $\underline{G}^0$ is itself a smooth group scheme, and may be identified with the Greenberg transform $Q^0 = \Gr^{R}_0(\underline{G}^0)$ \cite{cunningham-roe:13a}*{\S 4.3}. 
With $r\in \NN$ as above, let ${A}_r^0$ be the equivariant Weil sheaf on the algebraic group $G_r^i$ obtained by pull-back from $A^0$ along the affine morphism $G_r^i \to Q^0$.
Factor
\begin{equation}\label{eqn:tracerho}
\begin{tikzcd}
G^0(\Fq) \arrow{rr}{\trace(\orho^0)} \arrow{rd} && \EE \\
& G_r^0(\Fq) \arrow{ru}[swap]{\trace(\orho^0)_r} 
\end{tikzcd}
\end{equation}
Observe that $\trace(\orho^0)_r$ may be recovered from ${A}_m^0$:
\[
\trFrob{{A}^0_r} = \trace(\orho^0)_r
\]
